---
title: "Raw Data"
author: "DSPG 2019 ACPD Team"
date: "2019-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(echo = TRUE)
```

## Data Sources Description

### Arlington County Police Incident Log

Source: [Police Incident Log](http://bit.ly/21nPzoE)

Last Updated: 2019-07-10

Frequency: Daily

Time Coverage: 2015-01-01 - Present

Type: REST/JSON

### ABC Virginia Hearing and Appeals: Final Decisions

Source: [ABC Hearings and Appeals](https://www.abc.virginia.gov/enforcement/hearings-and-appeals/final-decisions/)

Last Updated: 2014-07 - 2018-06

Frequency: Annual

Type: Excel (XLS) One file per version

### ABC Virginia Licensee Data

Source: [ABC Licensees](https://www.abc.virginia.gov/licenses/licensee-search-staging)

Version 2010-04 - 2019-06

Type: Excel (XLSX)

### Arlington Restaurant Initiative Participants

Source: [Arlington Restaurant Initiative GIS Data](https://gisdata-arlgis.opendata.arcgis.com/datasets/arlington-restaurant-initiative)

Version: 2019-05-21

Type: GeoJSON / Feature Layer

### Arlington County, VA, 2010 Census Block Polygon boundaries.

Source: [Census Block 2010 Polygons](https://hub.arcgis.com/datasets/093fb97eabcc4829b72755a08c1b4766_0)

Version: Census 2010

Frequency: Decennial

Type: GeoJSON / Feature Layer

## Install Packages

```{r load packages}
if (!("pacman" %in% installed.packages()[,1])) {
  install.packages(pkgs = "pacman")
}
pacman::p_load(RPostgreSQL, tidyverse, janitor, maditr, httr, rio, sf)
```

## Environmental Variables Used

- `db_userid`
- `db_pwd`
- `arlington_api_key`

## Data Scraping

```{r police log}
crime_incidents <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    dbname = "acpd",
                    host = "postgis_1",
                    port = 5432L,
                    user = Sys.getenv("db_userid"),
                    password = Sys.getenv("db_pwd"))
  if (dbExistsTable(conn = conn, name = "police_log")) {
    police_log <- dbReadTable(conn = conn, name = "police_log")
    last_updated <- police_log$last_updated[1L]
  } else {
    last_updated <- today() - days()
  }
  if (last_updated < today()) {
    str_c("https://api.data.arlingtonva.us/api/v2/",
        "datastreams/POLIC-INCID-LOG-19589/data.csv/?auth_key=",
        Sys.getenv(x = "arlington_api_key")) %>%
      GET() %>%
      content(as = 'text', encoding = 'UTF-8') %>%
      fread() %>%
      dt_mutate(firstReportDtm = as_datetime(x = firstReportDtm, tz = "EST5EDT"),
                lastReportDtm = as_datetime(x = lastReportDtm, tz = "EST5EDT"),
                last_updated = today()) %>%
      setnames(old = 'reportNbr', new = 'id') %>%
      setnames(old = 'offenseDsc', new = 'description') %>%
      setnames(old = 'locationName', new = 'location') %>%
      setnames(old = 'latitudeCrd', new = 'latitude') %>%
      setnames(old = 'longitudeCrd', new = 'longitude') %>%
      setnames(old = 'firstReportDtm', new = 'start') %>%
      setnames(old = 'lastReportDtm', new = 'end') %>%
      dbWriteTable(con = conn,
                   name = 'police_log',
                   value = .,
                   row.names = FALSE,
                   overwrite = TRUE)
  }
  on.exit(expr = dbDisconnect(conn = conn))
}
crime_incidents()
```


```{r abc violations}
abc_violations <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    dbname = "acpd",
                    host = "postgis_1",
                    port = 5432L,
                    user = Sys.getenv("db_userid"),
                    password = Sys.getenv("db_pwd"))
  if (!dbExistsTable(conn = conn, name = "abc_violations")) {
    manual_fix <- function(x) {
      if (is.na(x = x)) {
        NA
      } else if (str_detect(string = x, pattern = "^\\d+$")) {
        x %>%
          as.integer() %>%
          excel_numeric_to_date() %>%
          as.character()
      } else {
        x %>%
          recode("05/12/216" = "2016-05-12",
                 "8/152014" = "2014-08-15",
                 "7-1613" = "2013-07-16",
                 "6/24/22017" = "2017-06-24",
                 "5/20/154" = "2015-05-20",
                 "10/9/185" = "2018-10-09",
                 "2/29/15" = "2015-03-01", # "2015-02-29" is not valid.
                 "6/31/16" = "2016-07-01", # "2016-06-31" is not valid.
                 "8/4/2014/" = "2014-08-04")
      }
    }
    map_df(.x = str_c("https://www.abc.virginia.gov/library/enforcement/other-documents/",
                      c("finalboardorders_july2017-june2018.xls",
                        "fy17-final-board-order-spreadsheet.xlsx",
                        "finalboardorders_july2015-june2016.xls",
                        "finalboardorders_july2014-june2015.xls"),
                      "?la=en"),
           .f = function(file) {
             x <- import(file = file,
                         col_types = c(rep(x = "text", times = 25L),
                                       rep(x = "skip", times = 7L)))
           }) %>%
      clean_names(case = "snake") %>%
      dt_filter(!is.na(state)) %>% # Excel is dirty. Valid records have state.
      dt_mutate(number_of_charges = as.integer(x = number_of_charges),
                violation_date = map_chr(.x = violation_date, .f = manual_fix) %>%
                  as_date(),
                repeat_abc_offense = repeat_abc_offense == "Y",
                continuance = continuance == "Y",
                date_available_to_h_a = map_chr(.x = date_available_to_h_a, .f = manual_fix) %>%
                  as_date(),
                date_heard = map_chr(.x = date_heard, .f = manual_fix) %>%
                  as_date(),
                decision_date = map_chr(.x = decision_date , .f = manual_fix) %>%
                  as_date(),
                disposition_date  = map_chr(.x = disposition_date , .f = manual_fix) %>%
                  as_date(),
                suspension_days = as.integer(x = suspension_days),
                civil_penalty = recode(.x = civil_penalty,
                                       "1.000.00" = "1000") %>%
                  as.integer(),
                mandatory_susp_w_civil_penalty_days =
                  as.integer(x = mandatory_susp_w_civil_penalty_days),
                probation_months = recode(.x = probation_months,
                                          "Y" = "12",
                                          "1 YR" = "12",
                                          "1 YEAR" = "12",
                                          "12 MOS" = "12"),
                revoked_y_n = revoked_y_n == "Y",
                dismissed_y_n = dismissed_y_n == "Y") %>%
      setnames(old = "revoked_y_n", new = "revoked") %>%
      setnames(old = "dismissed_y_n", new = "dismissed") %>%
      dt_arrange(abc_license_number) %>%
      dbWriteTable(conn = conn,
                   name = "abc_violations",
                   value = .,
                   row.names = FALSE)
  }
  on.exit(expr = dbDisconnect(conn = conn))
}
abc_violations()
```

```{r pull and create license table}
abc_licensees <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    dbname = "acpd",
                    host = "postgis_1",
                    port = 5432L,
                    user = Sys.getenv("db_userid"),
                    password = Sys.getenv("db_pwd"))
  if (!dbExistsTable(conn = conn, name = "abc_licensees")) {
    import(file = "https://www.abc.virginia.gov/library/licenses/other-documents/valicapril2019.xlsx?la=en") %>%
      clean_names(case = "snake") %>%
      dbWriteTable(con = conn,
                   name = 'abc_licensees',
                   value = .,
                   row.names = FALSE)
  }
  on.exit(expr = dbDisconnect(conn = conn))
}
abc_licensees()
```

```{r ari layer}
ari_restaurants_layer <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    host = "postgis_1",
                    dbname = "acpd",
                    port = 5432L,
                    user = Sys.getenv("db_userid"),
                    password = Sys.getenv("db_pwd"))
  if (!dbExistsTable(conn = conn, name = "ari_restaurants")) {
    st_read(dsn = "https://opendata.arcgis.com/datasets/2dc91350ea6f4bffb65ece8730b7fcca_0.geojson",
                 stringsAsFactors = FALSE) %>%
      clean_names(case = "snake") %>%
      data.table() %>%
      dt_mutate(ari_accredit = as_date(x = ari_accredit,
                                       format = "%m/%d/%Y",
                                       tz = "EST5EDT"),
                last_update = as_date(x = last_update,
                                      format = "%m/%d/%Y",
                                      tz = "EST5EDT"),
                accredit_date = as_date(x = accredit_date,
                                        format = "%m/%d/%Y",
                                        tz = "EST5EDT"),
                ask_angela = ask_angela == "Y") %>%
      st_write(dsn = conn, layer = "ari_restaurants")
  }
  on.exit(expr = dbDisconnect(conn = conn))
}
ari_restaurants_layer()
```

```{r census blocks}
census_blocks <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    host = "postgis_1",
                    dbname = "acpd",
                    port = 5432L,
                    user = Sys.getenv("db_userid"),
                    password = Sys.getenv("db_pwd"))
  if (!dbExistsTable(conn = conn, name = "arlington_census_blocks")) {
    st_read(dsn = "https://opendata.arcgis.com/datasets/093fb97eabcc4829b72755a08c1b4766_0.geojson") %>%
      clean_names(case = "snake") %>%
      st_write(dsn = conn, layer = "arlington_census_blocks")
  }
  on.exit(expr = dbDisconnect(conn = conn))
}
census_blocks()
```
