---
title: "R and SQL, Billboard Rap Charts"
output:
  html_document:
    df_print: paged
---


```{r}
rap_chart_artists <- readxl::read_excel("~/Rap_Charts.xlsx", sheet = "Artists")
rap_chart_singles <- readxl::read_excel("~/Rap_Charts.xlsx", sheet = "Singles")
```

## CREATE REUSABLE CONNECTION TO DATABASE
```{r warning=FALSE}
my_db_con <- sdalr::con_db("aschroed", pass=sdalr::get_my_password())
```

## WRITE DATA TO DATABASE TABLES
```{r}
DBI::dbWriteTable(my_db_con, "rap_chart_artists", rap_chart_artists, row.names=F)
DBI::dbWriteTable(my_db_con, "rap_chart_singles", rap_chart_singles, row.names=F)
```

## JOIN TABLES IN SQL WITH DBI
```{r}
DBI::dbGetQuery(my_db_con, 
                "SELECT * 
                 FROM rap_chart_artists a 
                 JOIN rap_chart_singles b ON a.\"Artist_ID\" = b.\"Artist_ID\"")
```

## JOIN TABLES IN SQL IN SQL CHUNK
```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
```

## JOIN TABLES IN SQL IN SQL CHUNK WITH OUTPUT VARIABLE
```{sql connection=my_db_con, output.var="rap_artist_singles"}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
```

```{r}
rap_artist_singles[, c("First_Name", "Last_Name")]
```

## JOIN TABLES IN SQL IN SQL CHUNK APPLY WHERE CLAUSE AND FUZZY SEARCH
```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
```

## JOIN TABLES IN SQL IN SQL CHUNK APPLY WHERE CLAUSE WITH 'AND' AND FUZZY SEARCH
```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
AND "Top_US_Rap_Chart" IS NOT NULL
```

## JOIN TABLES IN SQL IN SQL CHUNK APPLY SELECT AND WHERE CLAUSE WITH 'AND' AND FUZZY SEARCH AND ORDER
```{sql connection=my_db_con}
SELECT "Psuedonym", "Single", "Year", "Top_US_Rap_Chart"
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
AND "Top_US_Rap_Chart" IS NOT NULL
ORDER BY "Top_US_Rap_Chart", "Year"
```


## JOIN TABLES IN SQL IN SQL CHUNK APPLY SELECT AND WHERE CLAUSE AND FUZZY SEARCH AND ORDER AND GROUP BY WITH AGGREGATE FUNCTION
```{sql connection=my_db_con, max.print = 15}
SELECT "Psuedonym", "Top_US_Rap_Chart", COUNT("Top_US_Rap_Chart") chart_position_cnt
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%M.C.%'
AND "Top_US_Rap_Chart" IS NOT NULL
GROUP BY "Psuedonym", "Top_US_Rap_Chart"
ORDER BY "Psuedonym", "Top_US_Rap_Chart"
```



## SAME THING USING DATA.TABLE
```{r}
library(data.table)
rap_artist_singles_dt <- setDT(rap_artist_singles)
```

```{r}
rap_artist_singles_dt[Psuedonym %like% "LL"]
```

```{r}
rap_artist_singles_dt[Psuedonym %like% "LL", .(Psuedonym, Single, Top_US_Rap_Chart)]
```

```{r}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(Psuedonym, Single, Top_US_Rap_Chart)]
```

```{r}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(.N), c("Psuedonym", "Top_US_Rap_Chart")]
```

```{r}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(.N), c("Psuedonym", "Top_US_Rap_Chart")][order(Top_US_Rap_Chart)]
```

