<!--
Message to anyone editing this page
Note that most of the chunks are not being evaluated,
you need to pass in eval=TRUE into the chunk to evaluate the code chunk
-->
# SQL Databases

R4ds chapter (relational data): http://r4ds.had.co.nz/relational-data.html

Software-Carpentry SQL Lesson

- http://swcarpentry.github.io/sql-novice-survey/

DataCamp Courses

- https://www.datacamp.com/courses/intro-to-sql-for-data-science
- https://www.datacamp.com/courses/joining-data-in-postgresql

## Databases in R

```{r, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


```{r, eval = TRUE}
rap_chart_artists <- readxl::read_excel("data/Rap_Charts.xlsx", sheet = "Artists")
rap_chart_singles <- readxl::read_excel("data/Rap_Charts.xlsx", sheet = "Singles")
```

```{r, eval = TRUE}
head(rap_chart_artists)
```

```{r, eval = TRUE}
head(rap_chart_singles)
```

### CREATE REUSABLE CONNECTION TO DATABASE

```{r warning=FALSE, eval=FALSE}
# by default this will open up your own database (e.g., aschroed)
my_db_con <- sdalr::con_db(pass = sdalr::get_my_password())
```

### WRITE DATA TO DATABASE TABLES
```{r, eval=FALSE}
DBI::dbWriteTable(my_db_con, "rap_chart_artists", rap_chart_artists, row.names = FALSE)
DBI::dbWriteTable(my_db_con, "rap_chart_singles", rap_chart_singles, row.names = FALSE)
```

## Joins


### JOIN TABLES IN SQL WITH DBI
```{r, eval=FALSE}
DBI::dbGetQuery(my_db_con, 
                "SELECT * 
                 FROM rap_chart_artists a 
                 JOIN rap_chart_singles b ON a.\"Artist_ID\" = b.\"Artist_ID\"")
```

### JOIN TABLES IN SQL IN SQL CHUNK

<pre><code>```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
```</code></pre>

### JOIN TABLES IN SQL IN SQL CHUNK WITH OUTPUT VARIABLE

<pre><code>```{sql connection=my_db_con, output.var="rap_artist_singles"}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
```</code></pre>

```{r, eval=FALSE}
rap_artist_singles[, c("First_Name", "Last_Name")]
```

### JOIN TABLES IN SQL IN SQL CHUNK APPLY WHERE CLAUSE AND FUZZY SEARCH

<pre><code>```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
```</code></pre>

### JOIN TABLES IN SQL IN SQL CHUNK APPLY WHERE CLAUSE WITH 'AND' AND FUZZY SEARCH

<pre><code>```{sql connection=my_db_con}
SELECT *
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
AND "Top_US_Rap_Chart" IS NOT NULL
```</code></pre>

### JOIN TABLES IN SQL IN SQL CHUNK APPLY SELECT AND WHERE CLAUSE WITH 'AND' AND FUZZY SEARCH AND ORDER

<pre><code>```{sql connection=my_db_con}
SELECT "Psuedonym", "Single", "Year", "Top_US_Rap_Chart"
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%LL%'
AND "Top_US_Rap_Chart" IS NOT NULL
ORDER BY "Top_US_Rap_Chart", "Year"
```</code></pre>

### JOIN TABLES IN SQL IN SQL CHUNK APPLY SELECT AND WHERE CLAUSE AND FUZZY SEARCH AND ORDER AND GROUP BY WITH AGGREGATE FUNCTION

<pre><code>```{sql connection=my_db_con, max.print = 15}
SELECT "Psuedonym", "Top_US_Rap_Chart", COUNT("Top_US_Rap_Chart") chart_position_cnt
FROM rap_chart_artists a
JOIN rap_chart_singles s ON a."Artist_ID" = s."Artist_ID"
WHERE "Psuedonym" LIKE '%M.C.%'
AND "Top_US_Rap_Chart" IS NOT NULL
GROUP BY "Psuedonym", "Top_US_Rap_Chart"
ORDER BY "Psuedonym", "Top_US_Rap_Chart"
```</code></pre>


## dplyr and data.table

```{r, eval = TRUE}
library(dplyr)
library(data.table)

rap_chart_artists <- readxl::read_excel("data/Rap_Charts.xlsx", sheet = "Artists")
rap_chart_singles <- readxl::read_excel("data/Rap_Charts.xlsx", sheet = "Singles")

(rap_artist_singles <- rap_chart_artists %>%
    dplyr::full_join(rap_chart_singles, by = "Artist_ID"))
```

```{r, eval = TRUE}
rap_artist_singles_dt <- setDT(rap_artist_singles)
```

```{r, eval = TRUE}
rap_artist_singles_dt[Psuedonym %like% "LL"]
```

```{r, eval = TRUE}
rap_artist_singles_dt[Psuedonym %like% "LL", .(Psuedonym, Single, Top_US_Rap_Chart)]
```

```{r, eval = TRUE}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(Psuedonym, Single, Top_US_Rap_Chart)]
```

```{r, eval = TRUE}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(.N), c("Psuedonym", "Top_US_Rap_Chart")]
```

```{r, eval = TRUE}
rap_artist_singles_dt[Psuedonym %like% "LL" & !is.na(Top_US_Rap_Chart), .(.N), c("Psuedonym", "Top_US_Rap_Chart")][order(Top_US_Rap_Chart)]
```

