---
title: "Spatial R with SF Intro"
output:
  html_document:
    df_print: paged
---

# Code repository: git@devlab.vbi.vt.edu:sdal/sf_spatial_intro.git


## Install Latest ggplot
```{r}
#devtools::install_github("tidyverse/ggplot2")
```

## Load Libraries
```{r}
library(sf)
library(dplyr)
library(ggplot2) # latest github version
library(magrittr)
```

## Load Some Geography (in this case a shapefile)
```{r}
cd <- st_read("data/shapefiles/congressional_districts.shp", stringsAsFactors = FALSE)
head(cd)
```

## Plot the Geography
```{r}
plot(cd)
```

## Plot just One Column
```{r}
plot(cd[, "state"])
```


## Load Some Data (in this case, a csv file)
```{r}
cts <- read.csv('data/congressional_trump_scores.csv', stringsAsFactors = FALSE)[ , -1] %>%
  mutate(district = as.character(district))
head(cts)
```


## Combine Geography and Data
```{r}
dat <- left_join(cd, cts)
head(dat)
```


## Plot just One Column
```{r}
plot(dat[, "trump_score"])
```


## Fancy Plot with ggplot - first define a theme function (a set of layout/design parameters to re-use in each map)
```{r}
mapTheme <- function() {
  theme_void() + 
    theme(
      text = element_text(size = 7),
      plot.title = element_text(size = 11, color = "#1c5074", hjust = 0, vjust = 2, face = "bold"), 
      plot.subtitle = element_text(size = 8, color = "#3474A2", hjust = 0, vjust = 0),
      axis.ticks = element_blank(), 
      legend.direction = "vertical", 
      legend.position = "right",
      plot.margin = margin(1, 1, 1, 1, 'cm'),
      legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm")
    ) 
}
```


## Fancy Plot with ggplot - create ggplot and apply theme
```{r}
ggplot(dat) + 
  # plot a map with ggplot
  geom_sf(aes(fill = trump_score), color = NA) +
  # specify the projection to use
  coord_sf(crs = st_crs(102003)) +
  scale_fill_gradient2('Trump Score \n', low='#0099ff', mid = '#ffffff', high = '#ff6666', midpoint = 0.5) +
  labs(
    title = 'Where have U.S. Representatives voted with and against President Trump?',
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme()
```


## Filter to Smaller Data Set
```{r}
upper_mw <- dat %>%
  # select a few states using dplyr::filter
  filter(state %in% c('MN', 'IA', 'WI')) %>%
  # re-project to an appropriate coordinate system
  st_transform(2289) 
upper_mw_coords <- upper_mw %>%
  # find polygon centroids (sf points object)
  st_centroid %>%
  # extract the coordinates of these points as a matrix
  st_coordinates
# insert centroid long and lat fields as attributes of polygons
upper_mw$long <- upper_mw_coords[,1]
upper_mw$lat <- upper_mw_coords[,2]

head(upper_mw)
```


## Fancy Plot with ggplot - create ggplot and apply theme
```{r}
ggplot(upper_mw) + 
  # map districts by Trump Score
  geom_sf(aes(fill = trump_score), color = 'white') +
  # add labels according to locations of each polygon centroid
  geom_label(aes(long, lat, color = party, label = name), alpha = 0.75, size = 2) +
  scale_fill_gradient2('Trump Score \n', low='#0099ff', mid = '#ffffff', high = '#ff6666', midpoint = 0.5) +
  scale_color_manual('Political Party', values = c('Blue', 'Red')) +
  labs(
    title = "Congressional support for President Trump's policies in the Upper Midwest",
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme()
```


## Filter to a different Smaller Data Set
```{r}
by_state <- dat %>%
  group_by(state) %>%
  summarise(avg_trump_score = mean(na.omit(trump_score)), 
            districts = n_distinct(district))
head(by_state)
```


## Fancy Plot with ggplot - create ggplot and apply theme
```{r}
ggplot(by_state) + 
  geom_sf(aes(fill = avg_trump_score), color = 'white') +
  scale_fill_gradient2('Average \nTrump Score \nby State \n', low='#0099ff', mid = '#ffffff', high = '#ff6666', midpoint = 0.5) +
  coord_sf(crs = st_crs(102003)) +
  labs(
    title = "Which states' congressional delegations have supported President Trump's policies?",
    subtitle = "Mapping FiveThirtyEight's 'Trump Score' of House of Representative's voting records",
    caption = "Source: Azavea, Data: FiveThirtyEight"
  ) +
  mapTheme()
```


## Load Data from SDAL Database
```{r}
con <- sdalr::con_db("sdal", pass = sdalr::get_my_password())
evictions <- st_read(con, c("apps$dashboard", "va_bg_evictions_00_16"))
evictions <- evictions[evictions$item_geoparent=="Arlington County" & evictions$item_year=="2016",]

head(evictions)
```


## Change geometry column name and re-establish as the geometry column (BUG)
```{r}
colnames(evictions)[colnames(evictions)=="wkb_geometry"] <- "geometry"
st_geometry(evictions) <- "geometry"
```

```{r}
ggplot(evictions) + 
  geom_sf(aes(fill = item_value), color = 'white') +
  #scale_fill_gradient(low = "white", high = "black") +
  coord_sf(crs = st_crs(102003)) +
  labs(
    title = "Female Population by Block Group",
    subtitle = "Arlington County, VA",
    caption = "Source: American Community Survey"
  ) +
  mapTheme()
```


## Getting Data from a GIS API
### sf can read geographic files directly from a url! 

#### go to: https://hub.arcgis.com/pages/open-data
#### search for Hospitals
#### select from GeoPlatform.gov Source on the left side menu
#### select the first data set (should bring you to: https://hub.arcgis.com/datasets/geoplatform::hospitals-4)
#### from the API dropdown menu (on the right), copy the GeoJSON link
```{r}
hospitals <- st_read("https://opendata.arcgis.com/datasets/c56882e5caaa4ebd8e90a29adfaf24d0_0.geojson")
hospitals_va <- hospitals[hospitals$STATE=="VA",]
head(hospitals_va)
class(hospitals_va)
# plot(hospitals_va[,"TRAUMA"])
```

```{r}
ggplot(hospitals_va) + 
  geom_sf(aes(color = TRAUMA, size = POPULATION)) +
  #scale_fill_gradient(low = "white", high = "black") +
  coord_sf(crs = st_crs(102003)) +
  mapTheme()

# size bug
```



# Leaflet Example (Web Page Interactive Javascript)
## Prepare Data
```{r cache=TRUE, echo=FALSE, message=FALSE}
VABG <- readRDS("data/va_blockgroups_simplified_spdf.RDS")
cu12 <- readRDS("data/cu12.RDS")
cu12_cast <- data.table::dcast(cu12, NAMELSAD + ST_SUBST_FIPS ~ YEAR, value.var = c("VALUE", "MOE"))
map_data <- tmaptools::append_data(VABG, cu12_cast, key.shp = "GEOID", key.data = "ST_SUBST_FIPS")
map_data <- map_data[!is.na(map_data@data$NAMELSAD.data),]
max_value <- max(cu12$VALUE)
min_value <- min(cu12$VALUE)
```

## District Map (Census Block Groups)
```{r echo=FALSE, message=FALSE}
library(magrittr)

# pal <- leaflet::colorNumeric(
#   palette = "RdYlBu",
#   domain = min_value:max_value)

binpal <- leaflet::colorBin("viridis", min_value:max_value, 9, pretty = TRUE)

# labels <- sprintf("%s: %s", map_data_2015$NAMELSAD, map_data$VALUE)
years <- c("2015", "2014", "2013", "2012", "2011")
map <- leaflet::leaflet(width = "900px", data = map_data)
map <- leaflet::addProviderTiles(map = map, "CartoDB.Positron")
for (y in years) {
  col_name <- paste0("VALUE_", y)
  map <- leaflet::addPolygons(map = map,
                       smoothFactor = 0.2,
                       weight = 1,
                       fillOpacity = 0.7,
                       color = "white",
                       dashArray = "3",
                       fillColor = ~binpal(map_data@data[, c(col_name)]),
                       highlight = leaflet::highlightOptions(
                         weight = 5,
                         color = "#666",
                         dashArray = "",
                         fillOpacity = 0.7,
                         bringToFront = TRUE),
                       label = sprintf("Cnty: %s %s: %s", map_data$COUNTYFP, map_data$NAMELSAD, map_data@data[, c(col_name)]),
                       group = y
                       )
}
map <- leaflet::addLayersControl(map = map,
    baseGroups = years,
    position = "topleft",
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
map <- leaflet::addLegend(map = map, "topright", pal = binpal, values = ~map_data@data[, c(col_name)],
            title = "Households with Children < 12",
            opacity = 1
)
map
```


# To DataCamp! Spatial Analysis in R with sf and raster
# Go through Chapter 2: Plot vector and raster together