# Relational Data in R

Relational data chapter in r4ds

- http://r4ds.had.co.nz/relational-data.html

DataCamp Courses:

- https://www.datacamp.com/courses/joining-data-in-r-with-dplyr


Resources:

- [Rstudio dplyr Cheat Sheet](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
- dplyr join reference: https://dplyr.tidyverse.org/reference/join.html


## Create simple datasets

```{r}
library(tibble)
library(dplyr)
```

```{r}
x <- tribble(
    ~key, ~val_x,
    1, "x1",
    2, "x2",
    3, "x3"
)
x
```

```{r}
y <- tribble(
    ~key, ~val_y,
    1, "y1",
    2, "y2",
    4, "y3"
)
y
```

## one-to-one

When you perform a join, the function and code do not change.
The output of the join will be determined by whether or not you have duplicate values in your keys,
and whether or not both tables you are joining have duplicate values in the keys.

In a one-to-one merge, there are no duplicate keys in both tables.


```{r}
x %>% 
    inner_join(y, by = 'key')
```


```{r}
x %>% 
    left_join(y, by = 'key')
```

```{r}
x %>% 
    right_join(y, by = 'key')
```

```{r}
x %>% 
    full_join(y, by = 'key')
```

## many to one OR one to many

In a many-to-one merge, one of the tables have duplicate keys

```{r}
(x <- tribble(
    ~key, ~val_x,
    1, "x1",
    2, "x2",
    3, "x3",
    1, "x4",
    1, "x5",
    1, "x6"
))

(y <- tribble(
    ~key, ~val_y,
    1, "y1",
    2, "y2",
    4, "y3"
))
```

```{r}
x %>%
    left_join(y, by = 'key')
```

## many to many

In a many to many join, both tables will have duplicate keys in them.
You end up with what's known as a cartesian product of the keys.

```{r}
(x <- tribble(
    ~key, ~val_x,
    1, "x1",
    2, "x2",
    3, "x3",
    1, "x4",
    1, "x5",
    1, "x6"
))

(y <- tribble(
    ~key, ~val_y,
    1, "y1",
    2, "y2",
    4, "y3",
    1, "y4",
    1, "y5"
))
```

A many to many join creates a cartesian product

```{r}
x %>%
    left_join(y, by = 'key')
```



## Multiple keys

The below code won't work, but it's to show you that you can join by multiple keys

```r
x %>%
    left_join(y, by = c('key', 'col2', 'col3'))
```

## Different column names

You can also join tables when columns do not share the same name

```r
x %>%
    left_join(y, by = c('key' = 'key2', 'col2' = 'other', 'col3'))
```
