---
title: "Programmatically Mapping with R, sf and ggplot2"
author: "Adapted by Aaron Schroeder from the work of Mel Moreno and Mathieu Basille"
output:
  html_document:
    fig_caption: no
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    cache = TRUE,
    tidy = FALSE,
    fig.width = 7,
    fig.height = 7,
    out.width = "65%")
## This is just to "pre-load" all packages, and prevent the startup

library("ggplot2")
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("ggspatial")
library("maps")
library("tools")
library("ggrepel")
library("tmaptools")
```

## ggplot2

The package `ggplot2` implements the grammar of graphics in R, as a way to create code that make sense to the user: The grammar of graphics is a term used to breaks up graphs into semantic components, such as geometries and layers. Practically speaking, it allows (and forces!) the user to focus on graph elements at a higher level of abstraction, and how the data must be structured to achieve the expected outcome. While `ggplot2` is becoming the de facto standard for R graphs, it does not handle spatial data specifically. 

## sf
The current state-of-the-art of spatial objects in R relies on the Simple Features standard you see used in GIS systems in the package "sf".

Recently, the package `ggplot2` has allowed the use of simple features from the package `sf` as layers in a graph[^ggplotsf]. The combination of `ggplot2` and `sf` therefore enables to programmatically create maps, using the grammar of graphics, just as informative or visually appealing as traditional GIS software.

[^ggplotsf]: Note: Support of `sf` objects is available since version 3.0.0 of `ggplot2`.

## Getting Started
We start by loading the basic packages necessary for all maps, i.e. `ggplot2` and `sf`. We also suggest to use the classic dark-on-light theme for `ggplot2` (`theme_bw`), which is appropriate for maps:

```{r pkg-ggplot2-sf}
library("ggplot2")
theme_set(theme_bw())
library("sf")
```

The package `rnaturalearth` provides a map of countries of the entire world. Use `ne_countries` to pull country data and choose the scale (`rnaturalearthhires` is necessary for `scale = "large"`). The function can return `sp` classes (default) or directly `sf` classes, as defined in the argument `returnclass`:

```{r pkg-rnatural}
library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
head(world)
```


# General concepts illustrated with the world map

In this part, we will cover the fundamentals of mapping using `ggplot2` associated to `sf`, and presents the basics elements and parameters we can play with to prepare a map.


## Data and basic plot (`ggplot` and `geom_sf`)

First, let us start with creating a base map of the world using `ggplot2`. This base map will then be extended with different map elements, as well as zoomed in to an area of interest. We can check that the world map was properly retrieved and converted into an `sf` object, and plot it with `ggplot2`:

```{r world-plot, fig.width = 10, fig.height = 5, out.width = "90%"}
ggplot(data = world) +
    geom_sf()
```

This call nicely introduces the structure of a `ggplot` call: The first part `ggplot(data = world)` initiates the `ggplot` graph, and indicates that the main data is stored in the `world` object. The line ends up with a `+` sign, which indicates that the call is not complete yet, and each subsequent line correspond to another layer or scale. In this case, we use the `geom_sf` function, which simply adds a geometry stored in a `sf` object. By default, all geometry functions use the main data defined in `ggplot()`, but we will see later how to provide additional data.

Note that layers are added one at a time in a `ggplot` call, so the order of each layer is very important. All data will have to be in an `sf` format to be used by `ggplot2`; data in other formats (e.g. classes from `sp`) will be manually converted to `sf` classes if necessary.


## Title, subtitle, and axis labels (`ggtitle`, `xlab`, `ylab`)

A title and a subtitle can be added to the map using the function `ggtitle`, passing any valid character string (e.g. with quotation marks) as arguments. Axis names are absent by default on a map, but can be changed to something more suitable (e.g. "Longitude" and "Latitude"), depending on the map:
   
```{r world-title, fig.width = 10, fig.height = 5, out.width = "90%"}
ggplot(data = world) +
    geom_sf() +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("World map", subtitle = paste0("(", length(unique(world$sovereignt)), " countries)"))
```


## Map color (`geom_sf`)

In many ways, `sf` geometries are no different than regular geometries, and can be displayed with the same level of control on their attributes. Here is an example with the polygons of the countries filled with a green color (argument `fill`), using black for the outline of the countries (argument `color`):

```{r world-colors, fig.width = 10, fig.height = 5, out.width = "90%"}
ggplot(data = world) + 
    geom_sf(color = "black", fill = "lightgreen")
```

The package `ggplot2` allows the use of more complex color schemes, such as a gradient on one variable of the data. Here is another example that shows the population of each country. In this example, we use the "viridis" colorblind-friendly palette for the color gradient (with `option = "plasma"` for the plasma variant), using the square root of the population (which is stored in the variable `pop_est` of the `world` object):

```{r world-colors-pop, fig.width = 10, fig.height = 5, out.width = "90%"}
ggplot(data = world) +
    geom_sf(aes(fill = pop_est)) +
    scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```


## Projection and extent (`coord_sf`)

The function `coord_sf` allows to deal with the coordinate system, which includes both projection and extent of the map. By default, the map will use the coordinate system of the first layer that defines one (i.e. scanned in the order provided), or if none, fall back on WGS84 (latitude/longitude, the reference system used in GPS). Using the argument `crs`, it is possible to override this setting, and project on the fly to any projection. This can be achieved using any valid PROJ4 string (here, the European-centric ETRS89 Lambert Azimuthal Equal-Area projection):

```{r world-coord}
ggplot(data = world) +
    geom_sf() +
    coord_sf(crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")
```

Spatial Reference System Identifier (SRID) or an European Petroleum Survey Group (EPSG) code are available for the projection of interest, they can be used directly instead of the full PROJ4 string. The two following calls are equivalent for the ETRS89 Lambert Azimuthal Equal-Area projection, which is EPSG code 3035:

```{r world-coord-alt, eval = FALSE}
ggplot(data = world) +
    geom_sf() +
    coord_sf(crs = "+init=epsg:3035")

ggplot(data = world) +
    geom_sf() +
    coord_sf(crs = st_crs(3035))
```

The extent of the map can also be set in `coord_sf`, in practice allowing to "zoom" in the area of  interest, provided by limits on the x-axis (`xlim`), and on the y-axis (`ylim`). Note that the limits are automatically expanded by a fraction to ensure that data and axes don't overlap; it can also be turned off to exactly match the limits provided with `expand = FALSE`:

```{r world-zoom}
ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```


## Scale bar and North arrow (package `ggspatial`)

Several packages are available to create a scale bar on a map (e.g. `prettymapr`, `vcd`, `ggsn`, or `legendMap`). The package `ggspatial` provides some easy-to-use functions.

`scale_bar` that allows to add simultaneously the north symbol and a scale bar into the `ggplot` map. Five arguments need to be set manually: `lon`, `lat`, `distance_lon`, `distance_lat`, and `distance_legend`. The location of the scale bar has to be specified in longitude/latitude in the `lon` and `lat` arguments. The shaded distance inside the scale bar is controlled by the `distance_lon` argument. while its width is determined by `distance_lat`. Additionally, it is possible to change the font size for the legend of the scale bar (argument `legend_size`, which defaults to 3). The North arrow behind the "N" north symbol can also be adjusted for its length (`arrow_length`), its distance to the scale (`arrow_distance`), or the size the N north symbol itself (`arrow_north_size`, which defaults to 6). Note that all distances (`distance_lon`, `distance_lat`, `distance_legend`, `arrow_length`, `arrow_distance`) are set to `"km"` by default in `distance_unit`; they can also be set to nautical miles with "nm", or miles with "mi".


```{r world-scale-north, message = TRUE}
library("ggspatial")
ggplot(data = world) +
    geom_sf() +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97))
```

Note the warning of the inaccurate scale bar: since the map use unprojected data in longitude/latitude (WGS84) on an equidistant cylindrical projection (all meridians being parallel), length in (kilo)meters on the map directly depends mathematically on the degree of latitude. Plots of small regions or projected data will often allow for more accurate scale bars.


## Country names and other names (`geom_text` and `annotate`)

The `world` data set already contains country names and the coordinates of the centroid of each country (among more information). We can use this information to plot country names, using `world` as a regular `data.frame` in `ggplot2`. The function `geom_text` can be used to add a layer of text to a map using geographic coordinates. The function requires the data needed to enter the country names, which is the same data as the world map. Again, we have a very flexible control to adjust the text at will on many aspects:

* The size (argument `size`);
* The alignment, which is centered by default on the coordinates provided. The text can be adjusted horizontally or vertically using the arguments `hjust` and `vjust`, which can either be a number between 0 (right/bottom) and 1 (top/left) or a character ("left", "middle", "right", "bottom", "center", "top"). The text can also be offset horizontally or vertically with the argument `nudge_x` and `nudge_y`;
* The font of the text, for instance its color (argument `color`) or the type of font (`fontface`);
* The overlap of labels, using the argument `check_overlap`, which removes overlapping text. Alternatively, when there is a lot of overlapping labels, the package `ggrepel` provides a `geom_text_repel` function that moves label around so that they do not overlap. 
* For the text labels, we are defining the centroid of the counties with `st_centroid`, from the package `sf`. Then we combined the coordinates with the centroid, in the `geometry` of the spatial data frame. The package `sf` is necessary for the command `st_centroid`.

Additionally, the `annotate` function can be used to add a single character string at a specific location, as demonstrated here to add the Gulf of Mexico:

```{r world-country}
library("sf")
world_points <- st_centroid(world)
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))

ggplot(data = world) +
    geom_sf() +
    geom_text(data= world_points,aes(x=X, y=Y, label=name),
        color = "darkblue", fontface = "bold", check_overlap = FALSE) +
    annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
        fontface = "italic", color = "grey22", size = 6) +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```


# Final map

Now to make the final touches, the theme of the map can be edited to make it more appealing. We suggested the use of `theme_bw` for a standard theme, but there are many other themes that can be selected from (see for instance `?ggtheme` in `ggplot2`, or the package [`ggthemes`](https://cran.r-project.org/package=ggthemes) which provide several useful themes). Moreover, specific theme elements can be tweaked to get to the final outcome:

* Position of the legend: Although not used in this example, the argument `legend.position` allows to automatically place the legend at a specific location (e.g. `"topright"`, `"bottomleft"`, etc.);
* Grid lines (graticules) on the map: by using `panel.grid.major` and `panel.grid.minor`, grid lines can be adjusted. Here we set them to a gray color and dashed line type to clearly distinguish them from country borders lines;
* Map background: the argument `panel.background` can be used to color the background, which is the ocean essentially, with a light blue;
* Many more elements of a theme can be adjusted, which would be too long to cover here. We refer the reader to the documentation for the function `theme`.

```{r world-florida-final}
ggplot(data = world) +
    geom_sf(fill= "antiquewhite") +
    geom_text(data= world_points,aes(x=X, y=Y, label=name),
        color = "darkblue", fontface = "bold", check_overlap = FALSE) +
    annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
        fontface = "italic", color = "grey22", size = 6) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Map of the Gulf of Mexico and the Caribbean Sea") +
    theme(panel.grid.major = element_line(color = gray(.5),
        linetype = "dashed", size = 0.5),
        panel.background = element_rect(fill = "aliceblue"))
```


## Saving the map with `ggsave`

The final map now ready, it is very easy to save it using `ggsave`. This function allows a graphic (typically the last plot displayed) to be saved in a variety of formats, including the most common PNG (raster bitmap) and PDF (vector graphics), with control over the size and resolution of the outcome. For instance here, we save a PDF version of the map, which keeps the best quality, and a PNG version of it for web purposes:

```{r world-ggsave, eval = FALSE}
ggsave("map.pdf")
ggsave("map_web.png", width = 6, height = 6, dpi = "screen")
```


# Adding additional layers: an example with points and polygons

In the previous part, we presented general concepts with a map with little information (country borders only). The modular approach of `ggplot2` allows to successively add additional layers, for instance study sites or administrative delineations, as will be illustrated in this part.


## Field sites (point data)

We start by defining two study sites, according to their longitude and latitude, stored in a regular `data.frame`:

```{r sites-sites}
(sites <- data.frame(longitude = c(-80.144005, -80.109), latitude = c(26.479005, 
    26.83)))
```

The quickest way to add point coordinates is with the general-purpose function `geom_point`, which works on any X/Y coordinates, of regular data points (i.e. not geographic). As such, we can adjust all characteristics of points (e.g. color of the outline and the filling, shape, size, etc.), for all points, or using grouping from the data (i.e defining their "aesthetics"). In this example, we add the two points as diamonds (`shape = 23`), filled in dark red (`fill = "darkred"`) and of bigger size (`size = 4`):

```{r sites-plot}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sites, aes(x = longitude, y = latitude), size = 4, 
        shape = 23, fill = "darkred") +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

A better, more flexible alternative is to use the power of `sf`: Converting the data frame to a `sf` object allows to rely on `sf` to handle on the fly the coordinate system (both projection and extent), which can be very useful if the two objects (here world map, and sites) are not in the same projection. To achieve the same result, the projection (here WGS84, which is the CRS code #4326) has to be a priori defined in the `sf` object:

```{r sites-sf-create}
(sites <- st_as_sf(sites, coords = c("longitude", "latitude"), 
    crs = 4326, agr = "constant"))

(sites <- st_as_sf(sites, coords = c("longitude", "latitude"), 
    crs = 4326))
```

```{r sites-sf-plot}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

Note that `coord_sf` has to be called after all `geom_sf` calls, as to supersede any former input.


## States (polygon data)

It would be informative to add finer administrative information on top of the previous map, starting with state borders and names. The package `maps` (which is automatically installed and loaded with `ggplot2`) provides maps of the USA, with state and county borders, that can be retrieved and converted as `sf` objects:

```{r states}
library("maps")
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
head(states)
```

State names are part of this data, as the `ID` variable. A simple (but not necessarily optimal) way to add state name is to compute the centroid of each state polygon as the coordinates where to draw their names. Centroids are computed with the function `st_centroid`, their coordinates extracted with `st_coordinates`, both from the package `sf`, and attached to the state object:

```{r states-centroids}
(states <- cbind(states, st_coordinates(st_centroid(states))))
```

Note the warning, which basically says that centroid coordinates using longitude/latitude data (i.e. WGS84) are not exact, which is perfectly fine for our drawing purposes. State names, which are not capitalized in the data from `maps`, can be changed to title case using the function `toTitleCase` from the package `tools`:

```{r states-id}
library("tools")
states$ID <- toTitleCase(states$ID)
head(states)
```

To continue adding to the map, state data is directly plotted as an additional `sf` layer using `geom_sf`. In addition, state names will be added using `geom_text`, declaring coordinates on the X-axis and Y-axis, as well as the label (from `ID`), and a relatively big font size.

```{r states-plot}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) + 
    geom_text(data = states, aes(X, Y, label = ID), size = 5) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

We can move the state names slightly to be able to read better "South Carolina" and "Florida". For this, we create a new variable `nudge_y`, which is -1 for all states (moved slightly South), 0.5 for Florida (moved slightly North), and -1.5 for South Carolina (moved further South):

```{r states-nudge}
states$nudge_y <- -1
states$nudge_y[states$ID == "Florida"] <- 0.5
states$nudge_y[states$ID == "South Carolina"] <- -1.5
```

To improve readability, we also draw a rectangle behind the state name, using the function `geom_label` instead of `geom_text`, and plot the map again.

```{r states-plot-nudge}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) + 
    geom_label(data = states, aes(X, Y, label = ID), size = 5, fontface = "bold", 
        nudge_y = states$nudge_y) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```


## Counties (polygon data)

County data are also available from the package `maps`, and can be retrieved with the same approach as for state data. This time, only counties from Florida are retained, and we compute their area using `st_area` from the package `sf`:

```{r counties}
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
counties <- subset(counties, grepl("florida", counties$ID))
counties$area <- as.numeric(st_area(counties))
head(counties)
```

County lines can now be added in a very simple way, using a gray outline:

```{r counties-plot}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, fill = NA, color = gray(.5)) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

We can also fill in the county using their area to visually identify the largest counties. For this, we use the "viridis" colorblind-friendly palette, with some transparency:

```{r counties-plot-scale, fig.width = 9, out.width = "75%"}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, aes(fill = area)) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```


## Cities (point data)

To make a more complete map of Florida, main cities will be added to the map. We first prepare a data frame with the five largest cities in the state of Florida, and their geographic coordinates:

```{r cities-fl}
flcities <- data.frame(state = rep("Florida", 5), 
                       city = c("Miami", "Tampa", "Orlando", "Jacksonville", "Sarasota"), 
                       lat = c(25.7616798, 27.950575, 28.5383355, 30.3321838, 27.3364347), 
                       lng = c(-80.1917902, -82.4571776, -81.3792365, -81.655651, -82.5306527))
```

Instead of looking up coordinates manually, the package `tmaptools` provides a function `geocode_OSM`, which allows to retrieve geographic coordinates for any address, using the Open Street Map API. OSM, unlike the Google API, does not require an API key and is free! You can run the following code to automatically retrieve geographic coordinates of the five cities:

```{r cities-OSM}
library(tmaptools)
flcities <- geocode_OSM(c("Miami", "Tampa", "Orlando", "Jacksonville", "Sarasota"))
head(flcities)
```


We can now convert the data frame with coordinates to `sf` format:

```{r}
(flcities <- st_as_sf(flcities, coords = c("lon", "lat"), remove = FALSE, 
    crs = 4326, agr = "constant"))
```

We add both city locations and names on the map:

```{r cities-plot}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, fill = NA, color = gray(.5)) +
    geom_sf(data = flcities) +
    geom_text(data = flcities, aes(x = lon, y = lat, label = query), 
        size = 3.9, col = "black", fontface = "bold") +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

This is not really satisfactory, as the names overlap on the points, and they are not easy to read on the grey background. The package `ggrepel` offers a very flexible approach to deal with label placement (with `geom_text_repel` and `geom_label_repel`), including automated movement of labels in case of overlap. We use it here to "nudge" the labels away from land into the see, and connect them to the city locations:

```{r cities-plot-nudge}
library("ggrepel")
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, fill = NA, color = gray(.5)) +
    geom_sf(data = flcities) +
    geom_text_repel(data = flcities, aes(x = lon, y = lat, label = query), 
        fontface = "bold", nudge_x = c(1, -1.5, 2, 2, -1), nudge_y = c(0.25, 
            -0.25, 0.5, 0.5, -0.5)) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```


# Final map

For the final map, we put everything together, having a general background map based on the world map, with state and county delineations, state labels, main city names and locations, as well as a theme adjusted with titles, subtitles, axis labels, and a scale bar:

```{r layers-final-plot, fig.width = 9, out.width = "75%"}
library("ggspatial")
ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = counties, aes(fill = area)) +
    geom_sf(data = states, fill = NA) + 
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    geom_sf(data = flcities) +
    geom_text_repel(data = flcities, aes(x = lon, y = lat, label = query), 
        fontface = "bold", nudge_x = c(1, -1.5, 2, 2, -1), nudge_y = c(0.25, 
            -0.25, 0.5, 0.5, -0.5)) +
    geom_label(data = states, aes(X, Y, label = ID), size = 5, fontface = "bold", 
        nudge_y = states$nudge_y) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
    annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Observation Sites", subtitle = "(2 sites in Palm Beach County, Florida)") +
    theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"))
```

This example fully demonstrates that adding layers on `ggplot2` is relatively straightforward, as long as the data is properly stored in an `sf` object. Adding additional layers would simply follow the same logic, with additional calls to `geom_sf` at the right place in the `ggplot2` sequence.



## Getting Data from a GIS API
### sf can read geographic files directly from a url! 

#### go to: https://hub.arcgis.com/pages/open-data
#### search for Hospitals
#### select from GeoPlatform.gov Source on the left side menu
#### select the first data set (should bring you to: https://hub.arcgis.com/datasets/geoplatform::hospitals-4)
#### from the API dropdown menu (on the right), copy the GeoJSON link

```{r}
hospitals <- st_read("https://opendata.arcgis.com/datasets/c56882e5caaa4ebd8e90a29adfaf24d0_0.geojson")
hospitals_ia <- hospitals[hospitals$STATE=="IA",]
head(hospitals_ia)
class(hospitals_ia)
# plot(hospitals_va[,"TRAUMA"])
```

```{r}
ggplot(hospitals_ia) + 
  geom_sf(aes(color = TRAUMA, size = POPULATION)) +
  #scale_fill_gradient(low = "white", high = "black") +
  coord_sf(crs = st_crs(102003))
```