---
title: "Preprocessing of Data"
author: "DSPG 2019 ACPD Team"
date: "2019-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Housekeeping

```{r load packages}
if (!("pacman" %in% installed.packages()[,1])) {
  install.packages(pkgs = "pacman")
}
pacman::p_load(RPostgreSQL, tidyverse, maditr, httr, sf, geosphere, assertthat,
               measurements)
```

## Helpers

[MapQuest Geocoding API](https://developer.mapquest.com/documentation/geocoding-api/)

```{r geocode helper}
geocode <- function(address) {
  response <- str_c("http://www.mapquestapi.com/geocoding/v1/address?key=",
                    Sys.getenv(x = "mapquest_key"),
                    "&location=",
                    str_replace_all(string = address, pattern = "\\s+", replacement = "+") %>%
                      str_replace(pattern = "\\+Arlington,\\+VA\\+", replacement = "Arlington,VA")) %>%
    GET()
  assert_that(status_code(x = response) == 200L)
  response %>%
    content() %>%
    getElement(name = "results") %>%
    getElement(name = 1L) %>%
    getElement(name = "locations") %>%
    getElement(name = 1L) %>%
    getElement(name = "displayLatLng") %>%
    unlist() %>%
    t() %>%
    data.table()
}
```

Compute distance from the Clarendon Metro Station

```{r}
compute_distance_from_hq_in_mi <- function(lon, lat) {
  distm(x = c(lon, lat),
        y = c(-77.0950, 38.8872),
        fun = distVincentyEllipsoid) %>%
    conv_unit(from = "m", to = "mi")
}
```

Broad crime categories

```{r}
category_parser <- function(description) {
  if (str_detect(string = description,
                 pattern = "(?i)(assault|battery)")) {
    "Aggravated Assault"
  } else if (str_detect(string = description,
                        pattern = "(?i)(exposure|peace|disorder|urinat)")) {
    "Disorderly Conduct"
  } else if (str_detect(string = description,
                        pattern = "(?i)(drunk|intoxication|alcohol|liquor)")) {
    "Drunkenness"
  } else if (str_detect(string = description,
                        pattern = "(?i)(dui|container|breath|blood test|drivinig under influence|refusal of breath|drinking while driving)")) {
    "DUI"
  } else if (str_detect(string = description,
                        pattern = "(?i)(sex|rape|sodomy)")) {
    "Sexual Assault/Rape"
  } else if (str_detect(string = description,
                        pattern = "(?i)(minor|fake|fictitious|liquor misrepresenting age)")) {
    "Underage Drinking/Fake ID"
  } else if (str_detect(string = description,
                        pattern = "(?i)(parking violation|traffic violation)")) {
    "Traffic/Parking Violation"
    } else {
      "Other"
    }
  }
```

## Obtaining table with establishment (ABC license) and geolocation

- The ARI layer has the establishments and their geolocation

- Some establishments are missing in the layer

- Manually add the missing observations and geocode those with MapQuest

- Limit establishments to those that are ABC licensed

```{r establishments}
establishments <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    host = "postgis_1",
                    port = 5432L,
                    dbname = "acpd",
                    user = Sys.getenv(x = "db_userid"),
                    password = Sys.getenv(x = "db_pwd"))
  ari_restaurants <- st_read(dsn = conn, layer = "ari_restaurants") %>%
           data.table()
  licenses <- dbReadTable(conn = conn, name = "abc_licensees") %>%
           data.table()
  additional_data <- data.table(restaurant = c("Celtic House",
                                               "Lebanese Taverna-Pentagon Row",
                                               "Lebanese Taverna-Westover",
                                               "Fiona's Irish Pub",
                                               "G.O.A.T",
                                               "The Local Oyster",
                                               "Punch Bowl Social",
                                               "Rebellion On The Pike",
                                               "Wilson Hardware"),
                                address = c("2500 Columbia Pike, Arlington, VA 22204",
                                            "1101 S Joyce St, Arlington, VA 22202",
                                            "5900 Washington Blvd, Arlington, VA 22205",
                                            "567 23rd St S, Arlington, VA 22202",
                                            "3028 Wilson Blvd, Arlington, VA 22201",
                                            "4238 Wilson Blvd, Arlington, VA 22203",
                                            "4238 Wilson Blvd Ste. 1180, Arlington, VA 22203",
                                            "2900 Columbia Pike, Arlington, VA 22204",
                                            "2915 Wilson Blvd, Arlington, VA 22201"),
                                ari = c(rep(x = FALSE, times = 3L), rep(x = TRUE, times = 6L)),
                                ask_angela = NA) %>%
    cbind(map_df(.x = .$address, .f = geocode)) %>%
    st_as_sf(coords = c("lng", "lat")) %>%
    data.table()
  enhanced_ari <- rbind(ari_restaurants,
                        additional_data,
                        fill = TRUE) %>%
    dt_mutate(ari = if_else(condition = is.na(x = ari),
                            true = !is.na(x = ari_accredit),
                            false = ari))
  # Create common key
  enhanced_ari <- enhanced_ari %>%
    dt_mutate(key = paste(substr(tolower(restaurant),1,5), substr(tolower(address),1,5)) %>% 
                str_remove_all(pattern = "\\s"))
  licenses <- licenses %>%
    dt_mutate(key = paste(substr(tolower(trade_name),1,5), substr(tolower(add1),1,5)) %>%
                str_remove_all(pattern = "\\s"))
  DBI::dbSendQuery(conn, str_c("TRUNCATE TABLE ", "vabc_arlington_restaurants"))

  dbWriteTable(con = conn,
               name = 'vabc_arlington_restaurants',
               value = merge(x = enhanced_ari, y = licenses) %>%
                 st_as_sf() %>%
                 cbind(st_coordinates(.)) %>%
                 setDT() %>%
                 dt_filter(cty_cnty %in% "Arlington County") %>%
                 dt_select(-cty_cnty) %>%
                 setnames(old = "X", new = "x") %>%
                 setnames(old = "Y", new = "y"),
               row.names = FALSE,
               append = TRUE)
  on.exit(expr = dbDisconnect(conn = conn))
}
establishments()
```

## Crime description and geospatial and temporal filter to police interactions

```{r filtered incidents}
refresh_filtered_incidents <- function() {
  conn <- dbConnect(drv = PostgreSQL(),
                    host = "postgis_1",
                    port = 5432L,
                    dbname = "acpd",
                    user = Sys.getenv(x = "db_userid"),
                    password = Sys.getenv(x = "db_pwd"))

  DBI::dbSendQuery(conn, str_c("TRUNCATE TABLE ", "incidents_filtered"))

  dbReadTable(conn = conn, name = "police_log") %>%
    data.table() %>%
    dt_mutate(nearby = map2_dbl(.x = longitude,
                                .y = latitude,
                                .f = compute_distance_from_hq_in_mi) <= 0.5) %>%
    dt_mutate(day = weekdays(x = start, abbreviate = FALSE),
              hour = hour(x = start)) %>%
    # Broad alcohol-related
    dt_filter(str_detect(string = description,
                         pattern = "(DRUNK|DUI|ASSAULT|POSSESSION|SEXUAL|DISORDERLY)")) %>%
    # Broad crime categories
    dt_mutate(crime_category = map_chr(.x = description,
                                       .f = category_parser) %>%
                as.factor()) %>%
    dt_filter(crime_category != "Other") %>%
    # Create indicator for nightlife = THURS-SUN 21:00-05:00 (21:24, 1:5)
    # Also include Cinco de Mayo and St. Patricks Day
    dt_mutate(nightlife = (day %in% c("Thursday", "Friday", "Saturday", "Sunday") |
                             str_detect(string = start,
                                        pattern = "(05-05|05-06|03-17|03-18)$")) &
                (hour %in% c(1:5, 21:24))) %>%

    dbWriteTable(con = conn,
                 name = 'incidents_filtered',
                 value = .,
                 row.names = FALSE,
                 append = TRUE)
  on.exit(expr = dbDisconnect(conn = conn))
  }
refresh_filtered_incidents()
```
